## Project IDS##

1. Loading the file in a dataframe
```{r}
library(tidyverse)
dfhco<-data.frame(read_csv('C:\\SU\\IDS\\Project\\HMO_data.csv',show_col_types = FALSE))
```

2. Finding Missing values in Columns of the dataframe
```{r}
str(dfhco)
sum(is.na(dfhco$X))
sum(is.na(dfhco$age))
sum(is.na(dfhco$bmi))#78 missing values
sum(is.na(dfhco$children))
sum(is.na(dfhco$hypertension))#80 missing values
sum(is.na(dfhco$cost))
```

3. Clean up the NA's 
```{r}
library(imputeTS)
dfhco$bmi<-na_interpolation(dfhco$bmi)
dfhco$hypertension<-na_interpolation(dfhco$hypertension)
summary(dfhco)
```

4. Creating a **expensive** column based on cost 
```{r}
quantile(dfhco$cost)
#Expensive individuals will be those in the top 25% of the population which is cost greater than 4775 according to quantile function
dfhco$expensive <- with(dfhco, ifelse(cost > 4775, 'TRUE', 'FALSE'))
```
5. Plotting various Histogram for numeric variables
```{r}
hist(dfhco$age,main="Histogram of age")
hist(dfhco$bmi,main="Histogram of bmi")
hist(dfhco$children,main="Histogram of children")
hist(dfhco$hypertension,main="Histogram of hypertension")
hist(dfhco$cost,xlim=c(1,30000),main="Histogram of cost",breaks=20)
```
6. Creating Box Plots for expensive and inexpensive data with respect to cost and other variables 
```{r}
library(ggplot2)
Expensive <- filter(dfhco, cost > 4775)
Inexpensive <- filter(dfhco, cost <= 4775)
ggplot(Expensive)+aes(x=smoker,y=cost)+geom_boxplot()+ggtitle("Box Plot of Expensive data with respect to cost v/s smoker")
ggplot(Inexpensive)+aes(x=smoker,y=cost)+geom_boxplot()+ggtitle("Box Plot of Inexpensive data with respect to cost v/s smoker")
ggplot(Expensive)+aes(x=exercise,y=cost)+geom_boxplot()+ggtitle("Box Plot of Expensive data with respect to cost v/s excercise")
ggplot(Inexpensive)+aes(x=exercise,y=cost)+geom_boxplot()+ggtitle("Box Plot of Inexpensive data with respect to cost v/s excersie")
ggplot(dfhco)+aes(x=education_level,y=cost)+geom_boxplot()
ggplot(dfhco)+aes(x=yearly_physical,y=cost)+geom_boxplot()
```
7. Plotting Data on US Map with color shading as per cost  
```{r}
library(maps)
library(ggmap)
library(mapproj)
DF <- dfhco %>% group_by(location) %>% summarise(avgcost = mean(cost))
us<- map_data("state")
us$state_name <- tolower(us$region)
DF$location <- tolower(DF$location)
mappy <- merge(us,DF,by.x="state_name",by.y="location")
mappy<-mappy %>% arrange(order)
ggplot(mappy, aes(map_id= state_name))+aes(x=long, y=lat, group=group) +geom_polygon(aes(fill = avgcost), color = "black")+expand_limits(x=mappy$long, y=mappy$lat)+coord_map(projection = "mercator")
```
8. Using Association Rules to determine patterns i.e to get which other columns are more frequently appeared when cost was expensive

```{r}
library(arules)
dfhcon<-dfhco[,-14]
ruleset1<- apriori(dfhcon,
 parameter=list(supp=0.09, conf=0.3),
 control=list(verbose=F),
 appearance=list(default="lhs",rhs=("expensive=TRUE")))
inspect(ruleset1)
```
9. Creating Scatter Plot to determine distribution of data with respect to cost 

```{r}
ggplot(dfhco)+aes(x=bmi,y=cost,color=smoker)+geom_point()
ggplot(dfhco)+aes(x=bmi,y=cost,color=exercise)+geom_point() 
ggplot(dfhco)+aes(x=bmi,y=cost,color=yearly_physical)+geom_point()
ggplot(dfhco)+aes(x=bmi,y=cost,color=location_type)+geom_point()
```
10. Genrating Barplots
```{r}
ggplot(dfhco)+aes(x=education_level)+geom_bar(position="dodge",aes(fill=expensive))
ggplot(dfhco)+aes(x=children)+geom_bar(position="dodge",aes(fill=expensive))
ggplot(dfhco)+aes(x=gender)+geom_bar(position="dodge",aes(fill=expensive))
ggplot(dfhco)+aes(x=smoker)+geom_bar(position="dodge",aes(fill=expensive))
ggplot(dfhco)+aes(x=age)+geom_bar(position="dodge",aes(fill=expensive))
```

11. Training Model with SVM Against the dataset
```{r}
set.seed(111)
library(rio)
library(caret)
library(kernlab)
dfhco$expensive<-as.factor(dfhco$expensive)
trainList <- createDataPartition(dfhco$expensive, p=.7, list=FALSE)
trainSet <- dfhco[trainList,]
testSet <- dfhco[-trainList,]
svmmodel<- train(expensive ~ X+age+bmi+children+smoker+location+location_type+education_level+yearly_physical+exercise+married+hypertension+gender,data=trainSet , method= "svmRadial", trControl=trainControl(method = "none"), preProcess=c("center","scale"))
```

12. Use predict function to validate model against the test data
```{r}
svmPred<-predict(svmmodel,testSet,type='raw')
```

13. Creating Confusion Matrix.
```{r}
confusionMatrix(svmPred,testSet$expensive)
```

14. Training Model with SVM K-fold cross validation
```{r}
svmmodel.kfold<- train(expensive ~ X+age+bmi+children+smoker+location+location_type+education_level+yearly_physical+exercise+married+hypertension+gender,data=trainSet , method= "svmRadial", trControl=trainControl(method = "repeatedcv",number=10), preProcess=c("center","scale"))
```
15. Predicting on k-fold model
```{r}
svmPredkfold<-predict(svmmodel.kfold,testSet,type='raw')
```

16. Creating Confusion Matrix for k-fold.
```{r}
confusionMatrix(svmPredkfold,testSet$expensive)
```

17. Training rpart model
```{r}
library(e1071)
modelrpart<-train(expensive ~ X+age+bmi+children+smoker+location+location_type+education_level+yearly_physical+exercise+married+hypertension+gender, method = "rpart",data = trainSet)
```

18. Plotting Rpart Plot final model
```{r}
library(rpart.plot)
rpart.plot(modelrpart$finalModel)
prp(modelrpart$finalModel, faclen = 0, cex = 0.8, extra = 1)
plot(varImp(modelrpart))
```

19. Creating Confusion Matrix To calculate accuracy
```{r}
rpartpred<-predict(modelrpart,testSet,type='raw')
confusionMatrix(rpartpred,testSet$expensive)
```

20. Loading best model into rda file to save our model
```{r}
our_model <- modelrpart
save(our_model, file = "our_model.rda")
```